Contextual Combinatorial Cascading Bandits

### 2015-12-02 contextual combinatorial cascading bandits.tex
update main theorem and unify items

The above formulation is on disjuctive objective, that is, at a time step the agent stops as soon as she reveals a base arm at a position $k$ in the sequence with weight $1$ and she receives reward $\gamma_k$ as the result. Similarly, we could also consider the case of conjunctive objective. We also use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$ to indicate the weight of item $a$ at time $t$ satisfying equation (\ref{eq:expectation})). The learning agent observes all items until the first one with weight $0$. Suppose we also consider partial reward: if the $k$-th item is the first item with weight $0$, then the learning agent receives reward $\gamma_k'$; if all items have weight $1$, the learning agent receives reward $1$. The more items the learning agent reveals, the more reward the agent should receive. It is reasonable to assume that
$$
0 = \gamma'_1 \leq \gamma'_2 \leq \cdots \leq \gamma'_K \leq 1.
$$
At the end of time $t$, the learning agent observes $\bO_t, \bw_t(\ba_k^t), k \leq \bO_t$ and receives reward
\wei{For the following, why using $\gamma_{\bO_t - 1}'$ and $\gamma_{k - 1}'$ instead of $\gamma_{\bO_t}'$ and $\gamma_{k}'$. It is inconsistent with the above description, and I think the latter makes sense.} 
\shuai{This is an earlier version. I have modified it and all others.}
\begin{align*}
\br_{\wedge, t} &= \begin{cases}
\gamma_{\bO_t}'  &\text{if } \bw_t(\ba_{\bO_t}^t) = 0,\\
1 &\text{if } \bw_t(\ba_{\bO_t}^t) = 1,
\end{cases}\\
&=\begin{cases}
\gamma_{k}'  &\text{if } \bw_t(\ba_{i}^t) = 1, i < k, \bw_t(\ba_{k}^t) = 0,\\
1 &\text{if } \bw_t(\ba_{i}^t) = 1, i\leq \abs{\bA_t}.
\end{cases}
\end{align*}

@@

Let $\cH_t$ denote all history at the end of time $t$; $\cH_t$ consists of $\{x_s, \bA_s = (\ba_{1}^s,...,\ba_{\abs{\bA_s}}^s), \bO_s, \bw_s(\ba_k^s), k \in[\bO_s], s\in[t] \}$ and define $\EE_t[\cdot] = \EE[\cdot | \cH_t]$. Here we abuse the notation $\abs{\bA_s}$ to denote the number of base arms in the action $\bA_s$. By equation (\ref{eq:expectation}), we have 
$$
\EE[(\gamma_k \bw_{s,\ba_k^s}) | \cH_{s-1}] = \theta_*^{\top} (\gamma_k x_{s,\ba_k^s}).
$$
Using the ridge regression of data 
$$
\{(\gamma_k X_{s,\ba_k^s}, \gamma_k \bw_{s,\ba_k^s})\}_{k \in[\bO_s], s\in[t]},
$$
we get an $l^2$-regularized least-squares estimate of $\theta_*$ with regularization parameter $\lambda > 0$:
\begin{equation}
\hat{\theta}_t = (\bX_t^{\top}\bX_t + \lambda I)^{-1} \bX_t^{\top} \bY_t,
\end{equation}
where $\bX_t \in \RR^{(\sum_{s=1}^{t}\bO_s) \times d}$ is the matrix whose rows are $\gamma_k x_{s,\ba_k^s}^{\top}$ and $\bY_t$ is the column vector whose elements are $\gamma_k \bw_s(\ba_k^s)$, $k \in[\bO_s], s\in[t]$. Let
$$
\bV_t = \bX_t^{\top} \bX_t + \lambda I = \lambda I + \sum_{s=1}^{t} \sum_{k=1}^{\bO_s} \gamma_k^2 x_{s,\ba_k^s}x_{s,\ba_k^s}^{\top}.
$$
Then $\bV_t \in \RR^{d \times d}$ is a positive invertible matrix. The above notifications hold for both disjunctive and conjunctive objectives.

@@

Here is our main theorem.
\begin{theorem}
\label{thm:main}
Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \gamma_{K} \geq c f^*$ for some constant $c > 1$ and $\gamma_k' = 1 - \gamma_k, \gamma'_K \geq \frac{1}{4}f_{\wedge}^{\ast}$. Let regularization parameter satisfy $\lambda \geq C_\gamma$. Then for any $\delta > 0$, with probability at least $1 - \delta$, the cumulative expected regrets of our algorithm, {\em ConComCascade} for both conjunctive and disjunctive objectives, satisfy
\begin{align}
&R(T) = O(\frac{d}{c-1} \log(C_\gamma T) \sqrt{KT}),\\
&R_{\wedge}(T) = O(\frac{d}{f_{\wedge}^{\ast}} \log(C_\gamma T) \sqrt{KT}).
\end{align}
where
\begin{align*}
f^* &= \min_{1 \leq t \leq T} f_t^*\\
C_\gamma &= \sum_{k=1}^{K} \gamma_k'^2 = \sum_{k=1}^{K-1} (1 - \gamma_k)^2 = O(K).
\end{align*}
\end{theorem}

To prove theorem \ref{thm:main}, we could start from estimating $\EE[\Delta_{t, \bA_t}|\cH_t]$ for every time $t$. In fact, we can derive the following lemma.
\begin{lemma}
\label{lem:DeltaEstimate}
For any time $t$ and $A = (a_1, \ldots, a_{\abs{A}})$, if $f(A_t^*, \bU_t) \leq f(A, \bU_t)$, then we have
$$
\Delta_{t,A} \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}};
$$
if $f_{\wedge}(A_t^*, \bU_t) \leq f_{\wedge}(A, \bU_t)$, then we have
$$
\Delta_{t,A} \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}}.
$$
\end{lemma}

So if the chosen action $\bA_t$ is not suboptimal, which means
$$
f(A_t^*, \bU_t) \leq f(\bA_t, \bU_t).
$$
Then we can derive an estimate of $\Delta_{t, \bA_t}$, however, in terms of all elements in $\bA_t$.

However, this estimate involves all elements of $\bA_t$. Only when the learning agent has observed all elements of $\bA_t$, we can use such an estimate, which means we are able to estimate
$$
\EE[\Delta_{t, \bA_t}] = \EE[\Delta_{t, \bA_t}(1/p_{\bA_t}) \bOne\{\Delta_{t, \bA_t} > 0, \bO_t \geq \abs{\bA_t}\}].
$$
Problem might happen if $p_{\bA_t}$ is small. Thus we need the following lemma to find a good substitution of $\bA_t$.

\begin{lemma}
\label{lem:prefixExist}
Suppose the weights for items in $E$ are $w = (w_a)_{a \in E}$. Let $A = (a_1, ..., a_{\abs{A}})$ be a list and let $p_A$ denote the probability that all of items in $A$ are observed; equivalently, $p_A = \prod_{k=1}^{\abs{A}-1}w(a_i)$. Let $B_k = (a_1, ..., a_k), k \leq \abs{A}$ be a prefix of $A$. Then for any given constant $\frac{1}{2} < c < 1$, there exists a prefix $B$ of $A$ such that
$$
p_B \geq c f^* - \gamma_{\abs{B}}, \qquad \Delta_B \geq \frac{1}{2} \Delta_A,
$$
where
$$
f^* = \max_{A'} f(A', w), \qquad \Delta_{A'} = f^* - f(A', w).
$$
\end{lemma}

From this lemma, we can find a prefix $\bB_t$ of $\bA_t$ such that $p_{B_t}$ is large enough and $\Delta_{\bB_t}$ will not be too small. Then follow the above deductions, we can get a good estimate for $\EE[\Delta_{t, \bA_t}|\cH_{t-1}]$ in terms of $\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$. Once we have estimated $\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$ and $\beta_{t-1}(\delta)$, we can get an estimate for $\EE[\Delta_{t, \bA_t}|\cH_{t-1}]$, thus for $R(T)$. We leave all details in the next subsection.

>>>>

The above formulation is on disjuctive objective, that is, at a time step the agent stops as soon as she reveals a base arm at a position $k$ in the sequence with weight $1$ and she receives reward $\gamma_k$ as the result. Similarly, we could also consider the case of conjunctive objective. We also use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$ to indicate the weight of item $a$ at time $t$ satisfying equation (\ref{eq:expectation})). The learning agent observes all items until the first one with weight $0$. Suppose we also consider partial reward: if the $k$-th item is the first item with weight $0$, then the learning agent receives reward $\gamma_k'$; if all items have weight $1$, the learning agent receives reward $1$. The more items the learning agent reveals, the more reward the agent should receive. It is reasonable to assume that
$$
0 = \gamma'_1 \leq \gamma'_2 \leq \cdots \leq \gamma'_K \leq 1.
$$
At the end of time $t$, the learning agent observes $\bO_{\wedge, t}, \bw_t(\ba_k^t), k \leq \bO_{\wedge, t}$ and receives reward
\wei{For the following, why using $\gamma_{\bO_t - 1}'$ and $\gamma_{k - 1}'$ instead of $\gamma_{\bO_t}'$ and $\gamma_{k}'$. It is inconsistent with the above description, and I think the latter makes sense.} 
\shuai{This is an earlier version. I have modified it and all others.}
\begin{align*}
\br_{\wedge, t} &= \begin{cases}
\gamma_{\bO_{\wedge, t}}'  &\text{if } \bw_t(\ba_{\bO_{\wedge, t}}^t) = 0,\\
1 &\text{if } \bw_t(\ba_{\bO_{\wedge, t}}^t) = 1,
\end{cases}\\
&=\begin{cases}
\gamma_{k}'  &\text{if } \bw_t(\ba_{i}^t) = 1, i < k, \bw_t(\ba_{k}^t) = 0,\\
1 &\text{if } \bw_t(\ba_{i}^t) = 1, i\leq \abs{\bA_t}.
\end{cases}
\end{align*}

@@

Let $\cH_t$ denote all history at the end of time $t$; $\cH_t$ consists of $\{x_s, \bA_s = (\ba_{1}^s,...,\ba_{\abs{\bA_s}}^s), \bO_s, \bw_s(\ba_k^s), k \in[\bO_s], s\in[t] \}$ and define $\EE_t[\cdot] = \EE[\cdot | \cH_t]$. Here we abuse the notation $\abs{\bA_s}$ to denote the number of base arms in the action $\bA_s$. By equation (\ref{eq:expectation}), we have 
$$
\EE[(\gamma_k \bw_{s,\ba_k^s}) | \cH_{s-1}] = \theta_*^{\top} (\gamma_k x_{s,\ba_k^s}).
$$
Using the ridge regression of data 
$$
\{(\gamma_k X_{s,\ba_k^s}, \gamma_k \bw_{s,\ba_k^s})\}_{k \in[\bO_s], s\in[t]},
$$
we get an $l^2$-regularized least-squares estimate of $\theta_*$ with regularization parameter $\lambda > 0$:
\begin{equation}
\hat{\theta}_t = (\bX_t^{\top}\bX_t + \lambda I)^{-1} \bX_t^{\top} \bY_t,
\end{equation}
where $\bX_t \in \RR^{(\sum_{s=1}^{t}\bO_s) \times d}$ is the matrix whose rows are $\gamma_k x_{s,\ba_k^s}^{\top}$ and $\bY_t$ is the column vector whose elements are $\gamma_k \bw_s(\ba_k^s)$, $k \in[\bO_s], s\in[t]$. Let
$$
\bV_t = \bX_t^{\top} \bX_t + \lambda I = \lambda I + \sum_{s=1}^{t} \sum_{k=1}^{\bO_s} \gamma_k^2 x_{s,\ba_k^s}x_{s,\ba_k^s}^{\top}.
$$
Then $\bV_t \in \RR^{d \times d}$ is a positive invertible matrix. In the following, we will mainly use notations for disjunctive objective when the two cases are similar.

@@

Here is our main theorem.
\begin{theorem}
\label{thm:main}
Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_{K} \geq 0$ and $\gamma_k' = 1 - \gamma_k, \gamma'_K \geq \frac{1}{4}f_{\wedge}^{\ast}$. Let regularization parameter satisfy $\lambda \geq C_\gamma$. Then for any $\delta > 0$, with probability at least $1 - \delta$, the cumulative expected regrets of our algorithm, {\em ConComCascade} for both conjunctive and disjunctive objectives, satisfy
\begin{align}
&R(T) = O(\frac{d}{p^*} \log(C_\gamma T) \sqrt{KT}),\\
&R_{\wedge}(T) = O(\frac{d}{f_{\wedge}^{\ast}} \log(C_\gamma T) \sqrt{KT}).
\end{align}
where
\begin{align*}
&f^* = \min_{1 \leq t \leq T} f_t^{\ast},\\
&f_{\wedge}^* = \min_{1 \leq t \leq T} f_{\wedge, t}^{\ast},\\
&C_\gamma = \sum_{k=1}^{K} \gamma_k^2 = \sum_{k=1}^{K-1} (1 - \gamma_k')^2 = O(K),
\end{align*}
and $p_{t, \bA_t}$ and $p_{\wedge, t, \bA_t}$ are probabilities that all base arms of $\bA_t$ have been observed
\begin{align*}
&p_{t, \bA_t} = \PP(\bO_t \geq \abs{\bA_t}) = \prod_{k=1}^{\abs{\bA_t}-1} (1 - \theta_{\ast}^{\top} x_{t, \ba_k^t}),\\
&p_{\wedge, t, \bA_t} = \PP(\bO_{\wedge, t} \geq \abs{\bA_t}) = \prod_{k=1}^{\abs{\bA_t}-1} (\theta_{\ast}^{\top} x_{t, \ba_k^t}),\\
&p^* = \min_{1 \leq t \leq T} \min_{A \in \cS} ~ p_{t, A}.
\end{align*}
\end{theorem}

To prove theorem \ref{thm:main}, we could start from estimating $\EE[\Delta_{t, \bA_t}|\cH_t]$ for every time $t$. In fact, we can derive the following lemma.
\begin{lemma}
\label{lem:DeltaEstimate}
For any time $t$ and $A = (a_1, \ldots, a_{\abs{A}})$, if $f(A_t^*, \bU_t) \leq f(A, \bU_t)$, then we have
$$
\Delta_{t,A} \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}};
$$
if $f_{\wedge}(A_t^*, \bU_t) \leq f_{\wedge}(A, \bU_t)$, then we have
$$
\Delta_{\wedge, t,A} \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}}.
$$
\end{lemma}

So if the chosen action $\bA_t$ is not suboptimal, which means
$$
f(A_t^*, \bU_t) \leq f(\bA_t, \bU_t), ~~ f_{\wedge}(A_t^*, \bU_t) \leq f_{\wedge}(\bA_t, \bU_t).
$$
Then we can derive an estimate of $\Delta_{t, \bA_t}$ and $\Delta_{\wedge, t, \bA_t}$, however, in terms of all base arms in $\bA_t$. Only when the learning agent has observed all base arms of $\bA_t$, we can use such an estimate, which means we are able to estimate
\begin{align*}
&~~\EE_t[\Delta_{t, \bA_t}] \\
&= \EE[\Delta_{t, \bA_t}(1/p_{t, \bA_t}) \bOne\{\Delta_{t, \bA_t} > 0, \bO_t \geq \abs{\bA_t}\}],\\
&~~\EE_t[\Delta_{\wedge, t, \bA_t}] \\
&= \EE[\Delta_{\wedge, t, \bA_t}(1/p_{\wedge, t, \bA_t}) \bOne\{\Delta_{\wedge, t, \bA_t} > 0, \bO_{\wedge, t} \geq \abs{\bA_t}\}].
\end{align*}

In real problems involved with recommendations, the click rate is a very small number. Thus for the problem with disjunctive objective, $p^*$ is relatively high and can be regarded as a constant. So we have $p_{t, \bA_t} \geq p^*$. For the problem with conjunctive objective, $p_{\wedge, t, \bA_t}$ might be small. Thus we need the following lemma to find a good substitution of $\bA_t$.

\begin{lemma}
\label{lem:prefixExist}
Let $B_k = (\ba_1^t, \ldots, \ba_k^t), k \leq \abs{A}$ be a prefix of $\bA_t$. Then there exists a prefix $\bB_t$ of $\bA_t$ such that
$$
p_{\wedge, t, \bB_t} \geq \frac{1}{2} f_{\wedge, t}^* - \gamma_{\abs{\bB_t}}, \qquad \Delta_{\wedge, t, \bB_t} \geq \frac{1}{2} \Delta_{\wedge, t, \bA_t}.
$$
\end{lemma}
From this lemma, we can use the prefix $\bB_t$ to substitue $\bA_t$ because $p_{\wedge, t, \bB_t}$ is large enough and $\Delta_{\wedge, t, \bB_t}$ will not be too small. Then follow the above deductions, we can get a good estimate for $\EE[\Delta_{\wedge, t, \bA_t}|\cH_{t-1}]$ in terms of $\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$. Once we have estimated $\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$ and $\beta_{t-1}(\delta)$, we can get an estimate for $\EE[\Delta_{\wedge, t, \bA_t}|\cH_{t-1}]$, thus for $R_{\wedge}(T)$. We leave all details in the next subsection.
